---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}

df <- read.csv(unz("activity.zip","activity.csv"),header = TRUE, sep=",")
df$date <- as.Date(df$date,"%Y-%m-%d")

summary(df)
```



## What is mean total number of steps taken per day?
Removing the na values and making a histogram of the steps per day. 

```{r}
stepsperday <- df[!is.na(df$steps),]
stepsperday <- aggregate(. ~ date,stepsperday,sum)[,1:2]

hist(stepsperday$steps)
```


Mean and median of the total number of steps taken per day
```{r}
boxplot(stepsperday$steps,ylab="Number of steps per day")

```



## What is the average daily activity pattern?
```{r}
stepsperinterval <- aggregate(. ~ interval,df,mean)[,1:2]
plot(stepsperinterval$interval,stepsperinterval$steps, type="l", xlab="interval",
     ylab="average number of steps")

maxintervalvalue <- max(stepsperinterval$steps)
maxinterval <-  stepsperinterval[stepsperinterval$steps == maxintervalvalue,1]
```
The 5-minute interval, on average across all the days in the dataset, containing
the maximum number of steps is `r maxinterval` (`r maxintervalvalue`)


## Imputing missing values

```{r}
numberofmissingvalues <-  sum(is.na(df$steps))
```

There are `r numberofmissingvalues` rows in the dataset containing NAs. I will
replace these values by the mean of the dataset. Then I plot the histogram and the means and medians.

```{r}

 df2 <- transform(df, steps = ifelse(is.na(steps), mean(steps, na.rm=TRUE), steps))

stepsperday2 <- aggregate(. ~ date,df2,sum)[,1:2]

hist(stepsperday2$steps)
boxplot(stepsperday$steps,ylab="Number of steps per day (no NAs)")

```


We can see that these values do not differ a lot from the estimates from the first part of the assignment.
The total number of steps was `r sum(stepsperday$steps)` but is now `r as.integer(sum(stepsperday2$steps))`


## Are there differences in activity patterns between weekdays and weekends?


```{r}
df2$weekday <- factor(ifelse(weekdays(df2$date) %in% c("samedi","dimanche"), "weekend","weekday"))
stepsperdaytype <- aggregate(x= df2$steps,by=list(df2$interval,df2$weekday), FUN="mean")

library(ggplot2)
ggplot(stepsperdaytype,aes(x=x,y=Group.1,color=Group.2)) + geom_line() + xlab("Interval") + ylab("Average steps") + facet_grid(stepsperdaytype$Group.2 ~ .)


```



